---
title: "Get Historical Foreign Currency Rates from the Federal Reserve"
author: "Sharon Francisco"
date: "July 11, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Synopsis

The Federal Reserve posts historical foreign currency exchange rates for several currencies. This code gets the historical rates for British pounds (GBP).

The steps for extracting the data was obtained [here](http://statistics.berkeley.edu/computing/r-reading-webpages). 


## Installing Packages and Loading Libraries 
The packages used in the analysis are installed below.

```{r installAndLoadPackages}
if(!require(dplyr, quietly = TRUE)) install.packages("dplyr")
library(dplyr)
if(!require(xlsx, quietly = TRUE)) install.packages("xlsx")
library(xlsx)
```

## Set Root Directory and Load File


```{r setRootDir}
rootDir <- "C:/Users/b1050549/Documents/Anti_Corruption/0_ISC_payments/00_analysis_code"

knitr::opts_knit$set(root.dir = rootDir)
```

Download forex history from [federalreserve.gov](http://www.federalreserve.gov/releases/h10/hist/default.htm).

```{r downloadForexHistory}
gbpFileUrl <- "http://www.federalreserve.gov/releases/h10/hist/dat00_uk.htm"
euroFileUrl <- "http://www.federalreserve.gov/releases/h10/hist/dat00_eu.htm"

data <- readLines(gbpFileUrl)
head(data, 250)
```

After examining the html code, I found the HTML tags for the dates and rates in the table. I'll get the lines that have these tags.

```{r getDatesRatesHtml}
datePattern <- "<th id=\"r1\" headers=\"a1\">([^<]*)</th>"
ratePattern <- "<td headers=\"a2 a1 r1\">([^<]*)</td>"

dateLines <- grep(datePattern, data, value = TRUE)
head(dateLines, 25)

rateLines <- grep(ratePattern, data, value = TRUE)
head(rateLines, 25)
```

I'll take the html tags off the data.

```{r removeHTML}
dateHead <- "<th id=\"r1\" headers=\"a1\">"
dateEnd <- "</th>"
dates <- gsub(dateHead, "", dateLines)
dates <- gsub(dateEnd, "", dates)
head(dates, 25)

rateHead <- "<td headers=\"a2 a1 r1\">"
rateEnd <- "</td>"
rates <- gsub(rateHead, "", rateLines)
rates <- gsub(rateEnd, "", rates)
head(rates, 25)
```

Next, I'll remove the white/blank spaces in the data.

```{r removeWhiteSpaces}
dates <- trimws(dates)
rates <- trimws(rates)
head(dates, 25)
head(rates, 25)
```

I'll reformat the dates as dates.

```{r reformatDates}
dates <- strptime(dates, "%d-%b-%y")
class(dates)
head(dates, 25)
```


I'll convert the *ND* values in the rates data to the NA logical constant. Then I'll convert the vector to numeric.

```{r convertNDs}
ConvertND <- function(vector) {
        for(i in seq_along(vector)) {
                if(vector[i] == "ND") {
                        is.na(vector[i]) <- TRUE
                }
        }
        vector
}

rates <- ConvertND(rates)
rates <- as.numeric(rates)
head(rates, 25)
```

Now, I'll bind the columns and name them.

```{r bindColumns}
dates <- as.data.frame(dates)
rates <- as.data.frame(rates)
forex <- bind_cols(dates, rates)
names(forex) <- c("Date", "Rate")
forex
```


